# Peer-graded Assignment: Course Project 1
# Coursera Reproducible Research Week 2

The code assumes that the unzipped analysis file is present in your working directory.

## Loading and preprocessing the data
So let's read the data:
```{r results='hide'}
rawData <- read.csv("activity.csv")
```
## What is mean total number of steps taken per day?
Calculating the total number of steps taken per day:
```{r}
sum(rawData$steps, na.rm=T)
```
Let's create a histogram of the total number of steps taken each day:
```{r}
stepsPerDay <- aggregate(steps ~ date, rawData, sum)
hist(stepsPerDay$steps, xlab = "Number of steps", ylab = "Days", main="Histogram of steps taken per day")
```

Checking the mean and median number of steps taken per day:
```{r}
mean(stepsPerDay$steps)
median(stepsPerDay$steps)
```
## What is the average daily activity pattern?
Creating a time series plot of the 5-minute interval and the average number of steps taken:
```{r}
meanStepsPerInterval <- aggregate(steps ~ interval, rawData, mean)
plot(meanStepsPerInterval, type="l")
```

Which 5-minute interval contains the maximum average number of steps across all days?
```{r}
with(meanStepsPerInterval, meanStepsPerInterval[steps == max(steps),"interval"])
```
## Imputing missing values
How many missing values are there in our dataset?
```{r}
nrow(rawData[is.na(rawData$steps),])
```
Replacing the missing values with the average for the respective interval
```{r}
require(dplyr)
tmp <- rawData %>% group_by(interval) %>% mutate_each(funs(mean=mean(., na.rm=T)), steps)
imputedData = rawData
imputedData[is.na(imputedData$steps),"steps"] <- tmp[is.na(tmp$steps),"mean"]
```
Let's create a histogram of the total number of steps taken each day on the imputed data
```{r}
stepsPerDay <- aggregate(steps ~ date, imputedData, sum)
hist(stepsPerDay$steps, xlab = "Number of steps", ylab = "Days", main="Histogram of steps taken per day")
```

Checking the mean and median number of steps taken per day on imputed dataset:
```{r}
mean(stepsPerDay$steps)
median(stepsPerDay$steps)
```
Comparing these with the values calculated above on the original dataset, we can see that the medium has not change but the median has been increased a bit.

## Are there differences in activity patterns between weekdays and weekends?
Let's create a factor with the levels "weekend" and "weekday" based on the date column:
```{r}
imputedData$dayType <- ifelse(weekdays(as.Date(imputedData$date)) %in% c("Saturday", "Sunday"), "weekend", "weekday")
```
Finally we create a panel plot to compare the average number of steps taken on weekdays and weekends:
```{r}
require(lattice)
stepsPerDay <- aggregate(steps ~ dayType + interval, imputedData , mean)
xyplot(steps ~ interval | dayType, data = stepsPerDay, layout = c(1, 2), type="l", xlab = "Interval", ylab = "Number of steps")

```
